### 2012-03-25 paper_draft.tex
Added v2-chip numbers. Removed space in superscript page 2 paragraph 3 (we have to move footnotes to references anyway) Added dbSNP-citation to papers.bib and text

Since 2006, companies like 23andMe, deCODEme or FamilyTreeDNA have been offering Direct-To-Consumer (DTC) genetic testing. 
Those companies use DNA microarrays to screen for around 1 million SNPs spread over the human genome. In return, customers 
get an analysis of the results, as well as a file that includes the customer's raw individual genotypes. In 2011, 23andMe 
alone had over 100,000 customers\footnote{http://spittoon.23andme.com/2011/06/15/23andme-2011-state-of-the-database-address/} 
- the company recognizes the potential to perform GWAS with this amount of data by using surveys to ask their customers about 
traits and diseases. With the consent of the customer the data is used for association studies. 23andMe has published several 
articles in which it replicates known findings, but also finds new associations for Parkinson's Disease \cite{Eriksson2010, Do2011}. 
Over 30,000 23andme-customers have participated in those association studies.

Although companies like 23andMe are willing to use the data for scientific research,
it is not easy for individual scientists to access the data. 
The generation of biomedical data by private companies raises concerns about privacy \footnote{https://www.23andme.com/about/privacy/}, 
liabilty and consent \cite{Caulfield2011}. 
Nevertheless, individual customers are willingly sharing their data. Most do so by uploading their data to
their personal website or to open software repositories like \textit{GitHub}. 
This data is scattered and unorganized, making it hard to use this information. While projects like SNPedia try to keep track of all 
the genotyping files \cite{Cariaso2011}, they do not provide the information necessary to perform GWAS, as the phenotypic information is 
most often not attached to the genetic information. Projects that attach the phenotype to the genetic information, 
like the \textit{Personal Genome Project}, still do not allow for an easy re-use of the data, as they lack an advanced programming interface (API) 
or other methods by which researchers could download the data and not every customer of DTC genetic testing can participate in it.

@@

SNPs are catalogued by their unique identifier, which consists of a prefix (mostly \textit{rs}, rarely \textit{i}) and a unique number. This is a common format, 
which is employed by the NCBI dbSNP database and is also widely used and easily parsed from different literature sources. Publications from the different databases as 
well as the users' genotypes are associated with individual SNPs by the Rs-ID. Allele and genotype frequencies are updated regularly, based on the data present in openSNP.

>>>>

Since 2006, companies like 23andMe, deCODEme or FamilyTreeDNA have been offering Direct-To-Consumer (DTC) genetic testing. 
These companies use DNA microarrays to screen for around 0.5 to 1 million SNPs spread over the human genome. In return, customers 
get an analysis of the results, as well as a file that includes the customer's raw individual genotypes. In 2011, 23andMe 
alone had over 100,000 customers\footnote{http://spittoon.23andme.com/2011/06/15/23andme-2011-state-of-the-database-address/} 
- the company recognizes the potential to perform GWAS with this amount of data by using surveys to ask their customers about 
traits and diseases. With the consent of the customer the data is used for association studies. 23andMe has published several 
articles in which it replicates known findings, but also finds new associations for Parkinson's Disease \cite{Eriksson2010, Do2011}. 
Over 30,000 23andme-customers have participated in those association studies.

Although companies like 23andMe are willing to use the data for scientific research,
it is not easy for individual scientists to access the data. 
The generation of biomedical data by private companies raises concerns about privacy\footnote{https://www.23andme.com/about/privacy/}, 
liabilty and consent \cite{Caulfield2011}. 
Nevertheless, individual customers are willingly sharing their data. Most do so by uploading their data to
their personal website or to open software repositories like \textit{GitHub}. 
This data is scattered and unorganized, making it hard to use this information. While projects like SNPedia try to keep track of all 
the genotyping files \cite{Cariaso2011}, they do not provide the information necessary to perform GWAS, as the phenotypic information is 
most often not attached to the genetic information. Projects that attach the phenotype to the genetic information, 
like the \textit{Personal Genome Project}, still do not allow for an easy re-use of the data, as they lack an advanced programming interface (API) 
or other methods by which researchers could download the data and not every customer of DTC genetic testing can participate in it.

@@

SNPs are catalogued by their unique identifier, which consists of a prefix (mostly \textit{rs}, rarely \textit{i}) and a unique number. This is a common format, 
which is employed by the NCBI dbSNP database \cite{Sherry2001} and is also widely used and easily parsed from different literature sources. Publications from the different databases as 
well as the users' genotypes are associated with individual SNPs by the Rs-ID. Allele and genotype frequencies are updated regularly, based on the data present in openSNP.
